Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-07 paper.tex
New version of the abstract

A phenomenon that is often studied related to the human-like design of 
robots is \textit{anthropomorphism} -- people's tendency to perceive human-like 
characteristics in non-human agents. While this is a commonly discussed trait of 
human-robot interaction (HRI), the phenomenon paradoxically appears to lack formal 
grounds. Based on findings from several empirical studies and a 
comprehensive synthesis of literature, this article attempts at extending 
the current understanding of anthropomorphism. In accordance with theories 
from social psychology, cognitive science, and neuroscience, we suggest to 
formalize qualitative differences and dynamics over time in how 
humans anthropomorphize non-human agents. We propose several key factors that determine anthropomorphism, 
namely the 
design of the non-human agent, the individual characteristics of the human 
user, and the context of use. An initial model that we call 
\textit{``Dynamics of Anthropomorphism''} tries to formalize the qualitative changes in anthropomorphism over time. To 
account for these variances, our model describes the non-linear relationship between anthropomorphic effects of a human-robot interaction and the human user's growing experience with the system.
We also present a cognitive interpretation that considers 
these described dynamics. Further, we discuss scenarios and techniques to measure anthropomorphism. Finally, we make suggestions how others can 
apply our model to better describe the complexity in human's 
tendency to perceive human-like characteristics in non-human agents.

>>>>

\textit{Anthropomorphism} -- the people's tendency to perceive human-like
characteristics in non-human agents -- is a phenomenon that is most often
studied from the sole perspective of the human-like \emph{design} of robots.

By building upon experimental results from several studies, a comprehensive
synthesis of the literature, as well as insights from social psychology,
cognitive science, and neuroscience, this article proposes to extend our current
understanding of anthropomorphism into a \textbf{formal model} that defines a
\textbf{clear terminology}, accounts for the \textbf{non-monotonic dynamics} of
the phenomenon, and relate the observed \emph{effects} of anthropomorphism to
their \textbf{underlying cognitive precursors}.

We also examine the \textbf{factors} that determine anthropomorphism (the
design of the non-human agent, the individual psychological determinants of the
human user and the context of use) and we discuss several \textbf{episodes of
interaction} that influence the anthropomorphic projections on the robot, like
the \textbf{novelty effect}, the role of \textbf{familiarization} and the
impact of \textbf{disruptive behaviors}.

The article also discusses techniques and scenarios to measure anthropomorphism,
and we conclude with a few suggestions on how others can apply our model to
better describe our tendency to perceive human-like characteristics in non-human
agents.
